---
title: "data manipulation with dplyr"
output: github_document
---

```{r}
library (tidyverse)
```
 
options(tibble.print_min = 3)
```{r}
litters_data = read_csv("data_import_examples/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)
```

```{r}
pups_data = read_csv("data_import_examples/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```


#select

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

```{r}
select(litters_data, group:gd_of_birth)
```

```{r}
select(litters_data, -pups_survive)

```

#rename
```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

#use rename instead of select
```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

#select functions
there are certain functions that can be used: 'starts with(), ends with(), contains()

```{r}
select(litters_data, starts_with("gd") )
```

#to pull variables in front
```{r}
select(litters_data, litter_number, pups_survive, everything() )
```

##filter
will be using comparison operators ('>','>=','<','<=','==' and '!=')
'%in%' to detect if values appear in set, and 'in()' to find any missing values

logical comparisons 'True' or 'False'

```{r}

gd_of_birth == 20
pups_born_alive >= 2
pups_survive != 4
!(pups_survive == 4)
group %in% c("Con7", "Con8")
group == "Con7" & gd_of_birth == 20
```

```{r}
filter(litters_data, gd_of_birth == 20)
```

```{r}
drop_na(litters_data)
```
```{r}
drop_na(litters_data, wt_increase)
```

#mutate
creating new variable
```{r}
#litter_data2 
mutate(litters_data, wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group),
  # wait_gain_kg = wt_gain * 2.2
)
```

#arrange


```{r}
head(arrange(litters_data, group, pups_born_alive), 10)
```

reverse the order

```{r}
head(arrange(litters_data, desc(group), pups_born_alive), 10)
```

#piping

```{r}
litters_data_raw = read_csv("data_import_examples/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = 
  mutate(
    litters_data_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_data_with_vars_without_missing = 
  drop_na(litters_data_with_vars, wt_gain)
litters_data_with_vars_without_missing
```

another method

```{r}
litters_data_clean = 
  drop_na(
    mutate(
      select(
        janitor::clean_names(
          read_csv("data_import_examples/FAS_litters.csv", col_types = "ccddiiii")
          ), 
      -pups_survive
      ),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
    ),
  wt_gain
  )

litters_data_clean
```

third method:
```{r}
litters_data = 
  read_csv("data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_data
```

NOTE: '%>%' is "then": keyboard shortcuts: ctrl+shift+M


